Privacy-Respecting Type Error Telemetry at Scale

### 2023-12-09 programming23/paper.tex
acks
artifact checkpoint
error count ignore module switch data ... needs work to incorporate

% TODO BEFORE SUBMISSION

@@

The type error telemetry for a Roblox Studio session is gathered as a series of
records, all with the same per-session \panon{} session identifier.
This allows us to correlate telemetry across a single
session, but not between sessions.
In order to avoid swamping the telemetry servers, the client reports on a subset
of telemetry events via two levels of uniform random sampling:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions generate \emph{any} type error telemetry, and
  \item
    0.5\% of all type analyses (approximately 1 in 200 keystrokes)
    in an enrolled session generate a telemetry record.
\end{itemize}
In addition to sending telemetry for randomly-selected keystrokes,
our system sends a telemetry record every time the session changes
focus from one module to another.
These records provide important context for mode switches:
if one record uses \mstrict{} mode but the next uses \mnocheck{},
then it is critical to know whether the creator downgraded modes
or simply switched to another module.

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    \uid{}: a \panon{} identifier for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and \FS{} errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

@@

\Cref{f:error-density} plots changes to error density for every session in the dataset.
The x axis counts time in seconds from the start of the session.
Most sessions last a few dozen seconds~(\cref{t:session-size}), so the axis
ends at 150 seconds.
%% TODO explain why 150 ... shows some tail but not whole thing
The y axis shows changes to errors over time; a positive number means errors
increased since the last record, and a negative number represents a decrease.
While there are a few errors that go outside the plot bounds
(max \mnocheck{}: 101, max \mnonstrict{}: 3577, max \mstrict{}: 87), we focus
on $\pm{}50$ errors because most of the mass is in that region.
The points are color-coded to give each session the same color, but this is not
very useful because there are many more sessions than colors.

@@

  %% \item
  %%   TODO plot FS with a much bigger range ... tells a different story about buildup?

@@

  % TE: nocheck & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
  % TE: nonstrict & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
  % TE: strict & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  %
  % TE mod: nocheck & 20085 & [\pct{49.83}] & 906 & [\pct{2.25}] & 19317 & [\pct{47.92}] \\
  % TE mod: nonstrict & 13030 & [\pct{40.22}] & 5513 & [\pct{17.02}] & 13852 & [\pct{42.76}] \\
  % TE mod: strict & 419 & [\pct{39.31}] & 230 & [\pct{21.58}] & 417 & [\pct{39.12}] \\
  %
  % FS: nocheck & 270763 & [\pct{36.99}] & 238696 & [\pct{32.61}] & 222588 & [\pct{30.41}] \\
  % FS: nonstrict & 35330 & [\pct{37.28}] & 29483 & [\pct{31.11}] & 29955 & [\pct{31.61}] \\
  % FS: strict & 574 & [\pct{33.71}] & 561 & [\pct{32.94}] & 568 & [\pct{33.35}] \\
  %
  % FS mod: nocheck & 259865 & [\pct{39.09}] & 175643 & [\pct{26.42}] & 229271 & [\pct{34.49}] \\
  % FS mod: nonstrict & 33988 & [\pct{39.74}] & 20738 & [\pct{24.25}] & 30798 & [\pct{36.01}] \\
  % FS mod: strict & 488 & [\pct{36.53}] & 344 & [\pct{25.75}] & 504 & [\pct{37.72}] \\
  \begin{tabular}{l@{~~}lr@{}rr@{}rr@{}r}
    & & \zerowidth{\rbox{\addsym{}}} & & \zerowidth{\ybox{\keepsym{}}} & & \zerowidth{\gbox{\dropsym{}}} \\\midrule
    type error  & \mnocheck{}   & 20085 & [\pct{49.83}] & 906 & [\pct{2.25}] & 19317 & [\pct{47.92}] \\
    (in module) & \mnonstrict{} & 13030 & [\pct{40.22}] & 5513 & [\pct{17.02}] & 13852 & [\pct{42.76}] \\
                & \mstrict{}    & 419 & [\pct{39.31}] & 230 & [\pct{21.58}] & 417 & [\pct{39.12}] \\
    \\[-2ex]
    \FS         & \mnocheck{}   & 259865 & [\pct{39.09}] & 175643 & [\pct{26.42}] & 229271 & [\pct{34.49}] \\
    error       & \mnonstrict{} & 33988 & [\pct{39.74}] & 20738 & [\pct{24.25}] & 30798 & [\pct{36.01}] \\
    (in module) & \mstrict{}    & 488 & [\pct{36.53}] & 344 & [\pct{25.75}] & 504 & [\pct{37.72}] \\
  \end{tabular}
  \caption{How often do edits increase, maintain, or decrease the number of errors?}
  \label{f:error-changes}
\end{figure}

@@

%% -----------------------------------------------------------------------------
% \acks

% Thanks to Benjamin Chung for several helpful discussions about data analysis
% and effective plotting.
% Greenman was supported by
% NSF grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
% to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.

>>>>

% TODO BEFORE SUBMISSION
% - units on all plots

@@

The type error telemetry for a Roblox Studio session is gathered as a series of
records, all with the same  \panon{} session identifier.
This allows us to correlate telemetry across a single
session, but not between sessions.
In order to avoid swamping the telemetry servers, the client reports on a subset
of telemetry events via two levels of uniform random sampling:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions generate \emph{any} type error telemetry, and
  \item
    0.5\% of all type analyses (approximately 1 in 200 keystrokes)
    in an enrolled session generate a telemetry record.
\end{itemize}
In addition to sending telemetry for randomly-selected keystrokes,
our system sends a telemetry record every time the session changes
focus from one module to another.
These records provide important context for mode switches:
if one record uses \mstrict{} mode but the next uses \mnocheck{},
then it is critical to know whether the creator downgraded modes
or simply switched to another module.

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    \uid{}: a \panon{} identifier (16-digit random number) for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and \FS{} errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

@@

\Cref{f:error-density} plots changes to error density for every session in the dataset.
The x axis counts time in seconds from the start of the session.
Most sessions last a few dozen seconds~(\cref{t:session-size}), so the axis
ends at 150 seconds to show some but not all of the long tail.
The y axis shows changes to errors over time; a positive number means errors
increased since the last record, and a negative number represents a decrease.
While there are a few errors that go outside the plot bounds
(max \mnocheck{}: 101, max \mnonstrict{}: 3577, max \mstrict{}: 87), we focus
on $\pm{}50$ errors because most of the mass is in that region.
The points are color-coded by their session.
There are many colors througout the plot, which shows that the data comes
from many different sessions and is not dominated by a few high-traffic sessions.

@@

  \item
    When a color appears above the x axis, the same color usually appears below it as well.
    For example, \mnocheck{} has a cluster of purple points near the 20 second mark
    and $\pm 30$ density, which suggests that one session repeatedly added and removed errors.
    The next subsection conducts a thorough analysis.

@@

  %% with ALL data, modswitch + no switch
  % TE: nocheck & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
  % TE: nonstrict & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
  % TE: strict & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  %
  % TE mod: nocheck & 20085 & [\pct{49.83}] & 906 & [\pct{2.25}] & 19317 & [\pct{47.92}] \\
  % TE mod: nonstrict & 13030 & [\pct{40.22}] & 5513 & [\pct{17.02}] & 13852 & [\pct{42.76}] \\
  % TE mod: strict & 419 & [\pct{39.31}] & 230 & [\pct{21.58}] & 417 & [\pct{39.12}] \\
  %
  % FS: nocheck & 270763 & [\pct{36.99}] & 238696 & [\pct{32.61}] & 222588 & [\pct{30.41}] \\
  % FS: nonstrict & 35330 & [\pct{37.28}] & 29483 & [\pct{31.11}] & 29955 & [\pct{31.61}] \\
  % FS: strict & 574 & [\pct{33.71}] & 561 & [\pct{32.94}] & 568 & [\pct{33.35}] \\
  %
  % FS mod: nocheck & 259865 & [\pct{39.09}] & 175643 & [\pct{26.42}] & 229271 & [\pct{34.49}] \\
  % FS mod: nonstrict & 33988 & [\pct{39.74}] & 20738 & [\pct{24.25}] & 30798 & [\pct{36.01}] \\
  % FS mod: strict & 488 & [\pct{36.53}] & 344 & [\pct{25.75}] & 504 & [\pct{37.72}] \\

  %% with NON-moduleswitch data
% TE
%   nocheck & 961 & (\pct{38.50}) & 694 & [\pct{27.80}] & 841 & [\pct{33.69}] \\
% FS
%   nocheck & 156422 & (\pct{37.19}) & 153965 & [\pct{36.61}] & 110194 & [\pct{26.20}] \\
% TE mod
%   nocheck & 0 & (\pct{na}) & 0 & [\pct{na}] & 0 & [\pct{na}] \\
% FS mod
%   nocheck & 153913 & (\pct{36.96}) & 152735 & [\pct{36.68}] & 109807 & [\pct{26.37}] \\
% TE
%   nonstrict & 8623 & (\pct{37.67}) & 5263 & [\pct{22.99}] & 9004 & [\pct{39.34}] \\
% FS
%   nonstrict & 18927 & (\pct{37.41}) & 17791 & [\pct{35.17}] & 13872 & [\pct{27.42}] \\
% TE mod
%   nonstrict & 7185 & (\pct{36.61}) & 4768 & [\pct{24.30}] & 7672 & [\pct{39.09}] \\
% FS mod
%   nonstrict & 18573 & (\pct{37.09}) & 17604 & [\pct{35.15}] & 13905 & [\pct{27.76}] \\
% TE
%   strict & 442 & (\pct{38.27}) & 244 & [\pct{21.13}] & 469 & [\pct{40.61}] \\
% FS
%   strict & 296 & (\pct{31.66}) & 360 & [\pct{38.50}] & 279 & [\pct{29.84}] \\
% TE mod
%   strict & 251 & (\pct{35.35}) & 204 & [\pct{28.73}] & 255 & [\pct{35.92}] \\
% FS mod
%   strict & 272 & (\pct{32.04}) & 317 & [\pct{37.34}] & 260 & [\pct{30.62}] \\

  \begin{tabular}{l@{~~}lr@{}rr@{}rr@{}r}
    & & \zerowidth{\rbox{\addsym{}}} & & \zerowidth{\ybox{\keepsym{}}} & & \zerowidth{\gbox{\dropsym{}}} \\\midrule
    type error  & \mnocheck{}   & 20085 & [\pct{49.83}] & 906 & [\pct{2.25}] & 19317 & [\pct{47.92}] \\
    (in module) & \mnonstrict{} & 13030 & [\pct{40.22}] & 5513 & [\pct{17.02}] & 13852 & [\pct{42.76}] \\
                & \mstrict{}    & 419 & [\pct{39.31}] & 230 & [\pct{21.58}] & 417 & [\pct{39.12}] \\
    \\[-2ex]
    \FS         & \mnocheck{}   & 259865 & [\pct{39.09}] & 175643 & [\pct{26.42}] & 229271 & [\pct{34.49}] \\
    error       & \mnonstrict{} & 33988 & [\pct{39.74}] & 20738 & [\pct{24.25}] & 30798 & [\pct{36.01}] \\
    (in module) & \mstrict{}    & 488 & [\pct{36.53}] & 344 & [\pct{25.75}] & 504 & [\pct{37.72}] \\
  \end{tabular}
  \caption{How often do edits increase, maintain, or decrease the number of errors?}
  \label{f:error-changes}
\end{figure}

@@

%% -----------------------------------------------------------------------------
\acks

Thanks to Benjamin Chung for several helpful discussions about data analysis
and effective plotting.
Greenman was supported by
NSF grant \href{https://nsf.gov/awardsearch/showAward?AWD_ID=2030859&HistoricalAwards=false}{2030859}
to the CRA for the \href{https://cifellows2020.org}{CIFellows} project.
